Formalizing the Edmonds-Karp Algorithm

### 2016-03-02 paper/root.tex
modifed abstract intro

\maketitle
\begin{abstract}
We present a formalization of the Ford-Fulkerson method for computing the maximum flow in a network. 
The formal proofs are done in Isabelle/HOL. They closely follow a standard textbook proof and are relatively easy 
to read, even without being an Isabelle expert.

We then use stepwise refinement to obtain the Edmonds-Karp algorithm and prove its upper complexity bound of $O(|V||E|^2)$.
Further refinement yields a verified implementation, which is almost as fast as an unverified  
reference implementation in Java.

@@

This paper present a formalization of the correctness proof of the Ford-Fulkerson algorithm in Isabelle/HOL~\cite{NPW02}. Our proof is based on the informal proof of the algorithm which is presented in the book \emph{Introduction to algorithms}~\cite{CLRS09}. Due to practical importance of the algorithm, we also present a verified implementation of the algorithm. Isabelle/HOL provides some automation for generating the code corresponding to a verified algorithm, however, in order to use the code generating features, one needs to do the formalization with executability in mind. Being limited only to executable concepts in the formalization has the disadvantage of cluttering the proofs with implementation details. Such approach makes the proofs more complicated, and may even render proofs of medium complex algorithms unmanageable.

>>>>

\maketitle
\begin{abstract}
We present a formalization of the Ford-Fulkerson method for computing the maximum flow in a network.
Our formal proof closely follows a standard textbook proof, so it is easy to read even for those who 
are not expert in Isabelle/HOL--- the interactive theorem prover used for the formalization.

We then use stepwise refinement to obtain the Edmonds-Karp algorithm and we formally prove its termination.
Further refinement yields a verified implementation, which is almost as fast as an unverified reference implementation in Java.

%We present a formalization of the Ford-Fulkerson method for computing the maximum flow in a network. 
%The formal proofs are done in Isabelle/HOL. They closely follow a standard textbook proof and are relatively easy 
%to read, even without being an Isabelle expert.

%We then use stepwise refinement to obtain the Edmonds-Karp algorithm and prove its upper complexity bound of $O(|V||E|^2)$.
%Further refinement yields a verified implementation, which is almost as fast as an unverified  
%reference implementation in Java.

@@

This paper present a formalization of the correctness proof of the Ford-Fulkerson method in Isabelle/HOL~\cite{NPW02}. Our formalization is based on the informal proof presented in the book \emph{Introduction to algorithms}~\cite{CLRS09}. We then provide a verified implementation of the Edmons-Karp's algorithm~\cite{EK72}. This algorithm instantiates the Ford-Fulkerson method with breadth-first-search and provides a polynomial algorithm for solving the maximum flow problem in generic case--- maximum flow on networks with real edge capacities.

Isabelle/HOL provides some automation for generating the code corresponding to a verified algorithm, however, in order to use the code generating features, one needs to do the formalization with executability in mind. Being limited only to executable concepts in the formalization has the disadvantage of cluttering the proofs with implementation details. Such approach makes the proofs more complicated, and may even render proofs of medium complex algorithms unmanageable.
